---
title: "CProj2"
author: "Sneha Bhattacharjee"
date: "10/17/2020"
output:
  pdf_document: default
  html_document: default
---
### Statistical Inference Course Project 2
#Basic setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r global_options, include=FALSE}
library(knitr)
#Loading the required library
opts_chunk$set(fig.width=10, fig.height=8, warning=FALSE, message=FALSE)
```

## Brief Introduction
Load the ToothGrowth data and perform a few basic exploratory data analysis and thus :-
- Provide a basic summary of the data.
- Use confidence intervals and/or hypothesis tests to compare the tooth growth by supp and dose. 
(Use the techniques from class only, even if there are other approaches to this that are worth considering)
- State my conclusions and the assumptions needed for my conclusions that are inferred.

## Loading the Data
```{r}
# load the neccesary libraries
library(ggplot2)
library(datasets)
library(gridExtra)
library(GGally)

#To see the Effect of Vitamin C on Tooth Growth in Guinea Pigs
data(ToothGrowth)
toothg <- ToothGrowth 
toothg$dose <- as.factor(toothg$dose) #convert it to afactor
```


### Basic Summary of the data
```{r}

str(toothg)
summary(toothg)
head(toothg)
table(toothg$supp, toothg$dose)
```

```{r echo=FALSE}
p11 <- ggplot(data=toothg, aes(x=dose,y=len,fill=dose)) +
  geom_boxplot() + 
  theme(legend.position="none") + 
  facet_grid(.~supp) 
p21 <- ggplot(data=toothg, aes(x=supp,y=len,fill=supp)) +
  geom_boxplot() + 
  theme(legend.position="none") + 
  facet_grid(.~dose) 
p31 <- ggplot(data=toothg, aes(x=supp,y=len,fill=supp)) +
  geom_boxplot()
p41 <- ggplot(data=toothg, aes(x=dose,y=len,fill=dose)) +
  geom_boxplot()
grid.arrange(p11, p41, p21, p31, ncol = 2, nrow=2)
#ggpairs(toothg, 
#        upper = list(continuous = "density", combo = "box"),
#        lower = list(continuous="smooth"),        
#        colour = "dose",
#        axisLabels='show')
```

##To do some analysis based on Analysis of Variance (ANOVA)
```{r}

anout <- aov(len ~ supp * dose, data=toothg)
summary(anout)
```

# The results show that there is a notable interaction between the length (len) and dosage (dose) (F(1,54)=15.572;p<0.01)
Also a very clear effect on length(len) by supplement type (supp) (F(2,54)=92;p<0.01) is observed.
Lasly there is a minor interaction between the combination of supplement type (supp) and dosage (dose) compared to the length (len) (F(2,54)=4.107;p<0.05) observed.

```{r}
TukeyHSD(anout)
#The Tukey HSD  analysis shows that there are significant differences between each of the groups in supp and dose.

```

```{r}
confint(anout)
print(model.tables(anout,"means"),digits=3)
```

## Conclusions

It can be onferred that it is very clearly indicating that both the supplement and the dosage have clear independent effects on the length of the teeth of guinea pigs.
Moreover, those means on an average have longer teeth. Supplement type has a clear influence too, but OJ has a greater average teethgrowth in combination with dosages 0.5 and 1 than for the VC supplement.
While teeth length for the VC supplement vs the OJ in combiantion with dosage 2 has no significant effect (i.e., almost same mean & same confidence interval)

Thus,The fact remains however that these assumptions are based on the following facts:

*The guinea pigs are repesentative of the population of guinea pigs, 
* The dosage and supplement were randomly assigned and 
* The distribution of the means is normal.

This is my conclusion from the entire analysis of the given problem statement after observing all the simulations, distributions and calculations.
