---
title: "CProj1"
author: "Sneha Bhattacharjee"
date: "10/17/2020"
output:
  pdf_document: default
  html_document: default
---
### Introduction
In this peer graded assignment, i will be investigating the exponential distribution in R and hence compare it with the Central Limit Theorem. I will be investigating the distribution of averages of 40 exponentials. A total of 1000 simulations will be done.

### Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages and library
For the purpose of this analysis, i will be using the library 'ggplot2'.

```{r load-packages, message = FALSE}
library(ggplot2)
```

### Illustration of simulations
1. Show the sample mean and compare it to theoretical mean of distribution obtained.
As per given instructions, the exponential distributions can be simulated in R using rexp(n, lambda) where, lambda is rate parameter. The simulation is repeated 1000 times here. 
The theoretical mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda.
Here, for 1000 simulations, lambda is considered 0.2 and sample size ,i.e, n is 40.

```{r simu1.1}
#Setting the seed to ensure reproduceability
set.seed(11081990)

#Setting lambda as 0.2
lambda <- 0.2 
#Set values of n
n <- 40

#Obtain the sample mean of 1 simulation with the above parameters
#Obtain sample mean of 1000 simulations with the same above parameters
#Calculate the mean of this simulation of sample means
expoDist <- matrix(data=rexp(n * 1000, lambda), nrow=1000)
expoDistMeans <- data.frame(means=apply(expoDist, 1, mean))

#Plot the means obtained
ggplot(data = expoDistMeans, aes(x = means)) + 
  geom_histogram(binwidth=0.1) +  stat_function(fun = dnorm, color = "orange") +
  scale_x_continuous(breaks=round(seq(min(expoDistMeans$means), max(expoDistMeans$means), by=1)))
```

#Theoretical mean that is the centre of distribution is
```{r}
th <- 1/lambda
th  #Let Th be the expected mean 
```
```{r}
mean1000 <- mean(expoDistMeans$means)
mean1000
```

The centre of distribution of sample means of 40 exponentials is close to the theoretical center of the distribution.

### 2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
```{r simu1.2}
#Find the expected standard deviation 
sd <- 1/lambda/sqrt(n)
sd
#Variance of standard deviation obtained as
Vare <- sd^2
Vare
```
```{r}
sd_x <- sd(expoDistMeans$means)
sd_x
Var_x <- var(expoDistMeans$means)
Var_x
```

We can observe that the standar deviations are very close in value. Although minor differences exist as variance is the square of standard deviation.

### 3.To show that distribution is approximately normal.
```{r disb1.3}
#To plot the sample means obtained from the simulation.
ggplot(data = expoDistMeans, aes(x = means)) + 
  geom_histogram(binwidth=0.1, aes(y=..density..), alpha=0.2) + 
  stat_function(fun = dnorm, arg = list(mean = th , sd = sd), colour = "red", size=1) + 
  geom_vline(xintercept = th, size=1, colour="#CC0000") + 
  geom_density(colour="pink", size=1) +
  geom_vline(xintercept = mean1000, size=1, colour="#0000CC") + 
  scale_x_continuous(breaks=seq(th-3,th+3,1), limits=c(th-3,th+3))  + ggtitle("The Histogram for simulation of Samples Means with n = 1000") + scale_x_continuous("Sample Means") + ylab("Density")
```

### Conclusion and Inference
We know that a Normal Distribution follows a Bell-shaped curve, where the red line represents calculated normal distribution, which is comparable to the shape of histogram.
The Central Limit Theorem states that the sample means will become that of a standard normal distribution as the sample size increases as well as meeting two conditions of independence (n < 10% ) and normal, or if skewed distribution, then n > 30.
Thus, from the graph we see, the calculated distribution of means of random sapmled exponential distributions, overlaps quite properly with the normal distribution with the assumed value of lambda.

